Turtlebot SLAM

### 2013-01-20 trunk/Article/Article.tex
Visualization finished. Loop closing finished. Navigation evaluation finished.

\section{Introduction}
Nowadays a lot of tasks are performed by mobile robots, for example transportation, search and rescue and automated vacuum cleaning (figure \ref{fig:vacuum_cleaner}). To efficiently perform these tasks the robot needs a map of its environment. Sometimes the data about the environment is already available as a map. Using for example GPS the robot can then locate itself, making the planning and solving of the task at hand easier. But for tasks in unknown environments or where external referencing  like GPS is unavailable the robot should only rely on internal sensors and make its own map of the environment.

The acquisition of these maps has been a major research focus in the robotics community the last decades \cite{Grisetti}. Building or learning these maps under position uncertainty with inaccurate sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem. A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing.

Filtering approaches model the problem as an on-line state estimation. The state of the system consists in the current robot position and the map. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are usually referred to as on-line SLAM methods. The other type of approach estimates the full trajectory from the full set of measurements. These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques.

An intuitive way to formulate the SLAM problem is to cast it as a pose graph optimization problem. Where the nodes represent robot poses or landmarks and in which the edges encode sensor measurements that constraint the connected poses. Once such a graph is constructed the method will try to find a configuration of the nodes that is maximally consistent with the measurements.

The rest of this paper is structured as follows. Section 2 will provide the reader with the necessary background information on this topic. Section 3 is about the implementation details of the methods used. Section 4 will explain the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results. Section 5 discusses the acquired results of the experiments and proposes further research based on those. And the last section wraps up the paper with the conclusions of this research.

@@

\section{State of the art}
\subsection{ROS}
ROS (Robot Operating System) \cite{Quigley}, is the framework used in this paper to program the robots and it provides an environment to run simulations for the implemented algorithm. As described by Quigley et al. \cite{Quigley}, writing software for robots is difficult since there are so many possible hardware setups which makes code reuse nontrivial. Adding to this problem of hardware setups is the amount of code needed as it must contain a deep stack starting from driver-level software going up through perception, abstract reasoning and beyond. Since the required expertise is well beyond the capabilities of a single researcher, robotics software architectures must also support large scale software integration efforts.

@@

As time progresses and the robot explores more of its environment the map is constantly updated. This is programmed by saving laser scans for every pose of the robot, which, when traveling for a distance of $0.5$ meters is added and by that the occupancy grid is updated. Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.
Figure \ref{fig:random_map} shows a partial map created by the ROS gmapping package.
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/random_map.png}
	\caption{A partial map}
	\label{fig:random_map}
\end{figure}

\section{Experiments and Results}
\subsubsection{Benchmarks}

\section{Discussion}
The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.

>>>>

\section{Introduction}
Nowadays mobile robots are able to assist humans in a diverse set of problems. Transportation, search, rescue and automated vacuum cleaning are examples from different fields. Typically, mobile robots can move through an environment using legs or wheels. In addition they are able to gather information about their environment using sensors. Some even have devices similar to the human arm or hand which allows them to interact with the environment. The vaccuum cleaning robot presented in Figure \ref{fig:vacuum_cleaner} is an example of a robot which has whells to move around. Furthermore it is equipped with a set of sensors. These sensors provide data about the location of obstacles as well as, for example, gaps in the floor the robot is moving on. To efficiently perform this task robots needs a map of its environment. For some tasks map data is available for the robot from the beginning. Provided a map the robot can locate itself using the incoming sensor data. For tasks in unknown environments, however, the robot must rely on internal sensors and make its own map of the environment.

The acquisition of these maps has been a major research focus in the robotics community \cite{Grisetti} [\textbf{MORE SOURCES}]. Building or learning these maps under position uncertainty provided only sensor measurements is often referred to as the Simultaneous Localization And Mapping (SLAM) problem. A large variety of solutions is available to the SLAM problem, these approaches can be classified as either filtering or smoothing [\textbf{SOURCE}].

Filtering approaches model the problem as a so called ``on-line" state estimation problem. The state of such a system consists of the current robot position and the map. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods. The other type of approach estimates the full trajectory from the full set of measurements. These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques [\textbf{SOURCE}].

An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem. The nodes of the graph represent robot poses or landmarks. The edges encode sensor measurements that constraint the connected poses. Once such a graph is constructed the graph-based SLAM method tries to find a configuration of the nodes that is maximally consistent with the contraints contained in the edges. (\textbf{abstract graph picture here})

The rest of this paper is structured as follows. Section \ref{sec:sota} provides the reader with the necessary background information on the topic. Section \ref{sec:impl} is about the implementation details of the methods used. Section \ref{sec:exp} explains the experiments ran to validate and benchmark the implemented methods and also evaluate the obtained results. Section \ref{sec:disc} discusses the acquired results of the experiments and proposes further research based on those. And the last section wraps up the paper with the conclusions of this research.

@@

\section{State of the art}
\label{sec:sota}
\subsection{ROS}
ROS (Robot Operating System) \cite{Quigley}, is the framework used in this paper to program the robots and it provides an environment to run simulations for the implemented algorithm. As described by Quigley et al. \cite{Quigley}, writing software for robots is difficult since there are so many possible hardware setups which makes code reuse nontrivial. Adding to this problem of hardware setups is the amount of code needed as it must contain a deep stack starting from driver-level software going up through perception, abstract reasoning and beyond. Since the required expertise is well beyond the capabilities of a single researcher, robotics software architectures must also support large scale software integration efforts.

@@

As time progresses and the robot explores more of its environment the map is constantly updated. This is programmed by saving laser scans for every pose of the robot, which, when traveling for a distance of $0.5$ meters is added and by that the occupancy grid is updated. Upon a loop closure the occupancy grid is optimized with its information about obstacles and free spaces.

\section{Experiments and Results}
\label{sec:exp}
This section demonstrates the results of the Turtlebot SLAM implementation as described in section \ref{sec:impl}. Several outputs of the visualization tool RViz are presented and discussed. Comparisons are drawn between different variations of the implementation and the performance of the program is evaluated.

\subsection{Visualization}
When running the program via the launch file, RViz and Stage are automatically started. Stage shows the robot in its actual position on an accurate test map (see figure \ref{fig:stage_and_rviz} left). RViz shows the current belief of the robot about its environment, position and trajectory (see figure \ref{fig:stage_and_rviz} right).

The components of the pose graph and the environment are highlighted by the following labels:

\begin{description}
\item[green dots: ]past positions of the robot, saved as nodes of the pose graph
\item[red arrows: ]past orientations of the robot, included in the nodes of the pose graph
\item[blue lines: ]measurement constraints between the connected poses
\item[yellow dots: ]frontier points detected by the wavefront frontier detection algorithm
\item[orange dot: ] next navigation goal
\item[black area: ] detected obstacles
\item[light gray area: ] detected free space
\item[dark gray area: ] unexplored space
\end{description}

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/Stage_and_rviz.jpg}
	\caption{Robot and environment visualization with Stage (left) and RViz (right)}
	\label{fig:stage_and_rviz}
\end{figure}

\subsection{Loop closing effect}
One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected. This means that when the robot has been exploring an unknown area and then reenters an explored area, the transformation between the current pose and the past re-detected pose gives essential information on the complete pose graph, correcting the current belief about the environment and position.

The effect of loop closing detection is illustrated in figure \ref{fig:loop_closing_comparison}. The transparent red lines indicate the accurate map, while the underlying black areas indicate the belief of the robot about the obstacles of the environment.

The left part of figure \ref{fig:loop_closing_comparison} shows the explored map when walking an eight-shaped path through the map without applying loop detection. It can be seen that the belief about the map differs from the actual map in many areas, especially in the lower half of the map because this half has been explored last, exposing it to most of the measurement uncertainties.

The right part of figure \ref{fig:loop_closing_comparison} depicts the exploration when walking the same path as before but applying loop closing. This results in a pose graph with 7 more constrains, providing more information for correcting the poses and corresponding measurements. One can now see that the lower half of the belief about the map is much more accurate since the area in the center has been re-detected by the algorithm.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/Loop_closing_compared.jpg}
	\caption{Observed environment without (left) and with (right) loop closing of the pose graph}
	\label{fig:loop_closing_comparison}
\end{figure}

\subsection{Navigation evaluation}
An important aspect which has a high impact on the exploration behavior of the robot is the strategy by which the next navigation goal is determined. When facing the decision for the next goal, the given information are the detected frontiers, which can be spread all over the current map. The most obvious first implementation idea was to select the frontier which lies closest to the robot as the next navigation goal. This is definitely desirable because the robot can then explore this closest point very quickly and find out about the next frontiers.

Figure \ref{fig:navigation_comparison}(a) shows the resulting exploration path of the robot and the explored map after 300 seconds when applying this simple strategy. 10 nodes and 14 constraints have been created in the pose graph during this time which is very little information. Also the nodes are limited a small area of the map. This little degree of exploration is due to the fact that when only considering the closest frontier point there is a high risk that points are selected which are very close to obstacles. When the navigation algorithm tries to find a way to this point, it is inflating the obstacles first, making very close points unreachable. After a point has been discovered to be unreachable, the next closest point to the robot is chosen, which will likely be unreachable as well.

So it can be concluded that the number of close obstacles have to be taken into account in order to select reachable points as next goals. In the upcoming strategy, the sum of obstacles within a radius of 50 grid cells is added to the distance to the robot for every frontier point. The point with the smallest sum is then selected as the next goal. The application of this strategy is shown in figure \ref{fig:navigation_comparison}(b). The robot managed to explore the map within 300 seconds, creating a pose graph with 104 nodes and 150 constraints. However, one can see that the robot did not take a very exploration-efficient route through the map. There are a lot of unreasonable turns in areas that are already explored, while unexplored areas get left out for the time being.

Another aspect is to be added to the evaluation of every frontier point. That is the distance between the currently evaluated frontier point and the previous navigation goal. When this distance is minimized additionally to the distance to the robot and the number of close obstacles, the algorithm will select the next goals in a chain-like fashion, creating a rather intuitive exploration path through the map (figure \ref{fig:navigation_comparison}(c)). The robot was able to explore the map within only 150 seconds, creating a much smaller pose graph of 64 nodes and 66 constraints. This evaluation feature has been given a weight factor of 5 to increase the effect on the goal selection.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/Navigation_comparison.jpg}
	\caption{Resulting robot trajectory when applying different strategies for selecting the next navigation goal}
	\label{fig:navigation_comparison}
\end{figure}

\subsection{Runtime}

\section{Discussion}
\label{sec:disc}
The Wavefront Frontier detection algorithm \citep{Keidar} succeeds in finding frontiers in a fast and accurate manner.
